---
title: "loops and functions"
author: "Guillermo Romero"
date: "2022-08-11"
output: html_document
editor_options: 
  chunk_output_type: console
---

```{r}
library(tidyverse)
```

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

Create two sequences, one called weekdays that contains days of the week ("Monday", "Tuesday", "Wednesday", etc.) and one called transects that contains the series of transect names "Transect A", "Transect B,"Transect C".

```{r}
weekdays <- c("Monday","Tuesday", "Wednesday","Thursday", "Friday", "Saturday", "Sunday")

transects <- c("Transect A", "Transect B", "Transect C")

empty_matrix <- matrix(nrow = length(weekdays), ncol = length(transects))
# Write A nested For loop


for (i in seq_along(weekdays)){
  for (j in seq_along(transects)){
    day_transect <- paste(weekdays[i], "-", transects[j])
    empty_matrix[i,j] <- day_transect
  }
}
empty_matrix
```

Write a function called force that calculates a force (in Newtons), given inputs of mass (in kg) and acceleration (in ms2 (recall: F=ma), and returns a statement "The resulting force is \_\_\_ Newtons."

```{r}
force <- function(mass, acceleration){
  Force <- mass * acceleration
  print(paste("The resulsting force is", Force, "Newtons."))
}

force(10,10)
```

# Task 3

The length:weight relationship for fish is: W=aL\^b, where where L is total fish length (centimeters), W is the expected fish weight (grams), and a and b are species-dependent parameter values

| sci_name            | common_name     |  a_est | b_est |
|:--------------------|:----------------|-------:|------:|
| Chanos chanos       | Milkfish        | 0.0905 |  2.52 |
| Sphyraena barracuda | Great barracuda | 0.0181 |  3.27 |
| Caranx ignobilis    | Giant trevally  | 0.0353 |  3.05 |

Recreate the table above as a data frame stored as `fish_parms`. Then, write a function called `fish_weight` that allows a user to only enter the *common name* (argument `fish_name`) and *total length* (argument `tot_length`) (in centimeters) of a fish, to return the expected fish weight in grams. Test it out for different species and lengths.

Now, try creating a vector of lengths (e.g.0 to 100, by increments of 1) and ensuring that your function will calculate the fish weight over a *range* of lengths for the given species (try this for milkfish, storing the output weights as `milkfish_weights`.



```{r}
fish_parms <- tribble(
  ~sci_name, ~common_name, ~a_est, ~b_est, 
  "Chanos chanos", "Milkfish", 0.0905,2.52, 
  "Sphyraena barracuda", "Great barracuda", 0.0181, 3.27, 
  "Caranx ignobilis", "Giant trevally",0.0353, 3.05)

fish_weight <- function(fish_name, tot_length){
    c_name <- filter(fish_parms, common_name == fish_name)
    a_est <- c_name$a_est
    b_est <- c_name$b_est
    weight <- a_est * tot_length ^ b_est
    return(weight)}

fish_weight("Milkfish", 10)

zero_100 <- seq(0,100,1)

milkfish_weights <- fish_weight("Milkfish",zero_100)
milkfish_weights
```

# Task 4
Wave power (more accurately wave energy flux) in deep water is approximated by:

$$P_deep=0.5H2T$$
where P is power in kWm (potential power per wave meter), H is wave height in meters (more specifically, the significant wave height), and T is the wave period in seconds. Learn more here.

Write a function called wave_power that calculates potential ocean wave power given inputs of wave height and period.

Use your wave_power function to approximate wave power for a period of 8 seconds, over a range of wave heights from 0 to 3 meters by increments of 0.2 meters.

```{r}
wave_power <- function(wave_height,period){
  power_wave <- 0.5 * wave_height^2 * period
  return(power_wave)
}

wave_power(seq(0,3,0.2),8)
```

